---
title: "HW4"
author: "Sean Reyes"
date: "2024-03-20"
output: html_document
---

#### UTEID: str786
#### Github link:

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readr)
library(ggplot2)
library(mosaic)

gasprices <- read_csv("Downloads/Statistical Thinking/gasprices.csv")

```

#### Claim A:
Gas stations charge more if there is no direct competition in sight. 


#### Evidence:
```{r echo=FALSE}
# Theory A: found mean by whether or not competitors nearby, then bootstrapped for confidence interval of mean
mean(Price ~ Competitors, data = gasprices)

boot_pricecomp = do(10000)*mean(Price ~ Competitors, data=mosaic :: resample(gasprices))

confint(boot_pricecomp, level = .95)
```

Based on the statistical calculations, the gas prices of stations with no competition in sight had a mean price that fell between 1.85 and 1.90 with a 95% confidence (rounded to the nearest hundredth), while those that did have competition in sight had a mean price between 1.83 and 1.87 with 95% confidence (rounded to the nearest hundredth). The effect of this theory is not that large since gas stations without competition in sight on average tend to charge 2 to 3 cents more with 95% confidence. After re-sampling numerous times, the average prices were 1.88 +-0.2 dollars per gallon for no competition, and 1.85 +-0.2 dollars per gallon with competition, and this difference in mean is also shown by this box plot:

```{r echo=FALSE}
#boxplot comparing medians of prices faceted by competitors=yes/no
price_competitors = gasprices%>%
  select(Price, Competitors)

ggplot(price_competitors)+
  geom_boxplot(aes(y=Price))+
  facet_grid(~Competitors)
```

#### Conclusion: 
The data supports the claim that gas stations charge more when there are no competition in sight by about 2-3 cents per gallon with 95% confidence.

\

#### Claim B: 
Gas prices are more expensive in a richer area of the city.

```{r echo=FALSE}
income_price = gasprices%>%
  select(Price, Income)%>%
  arrange(Income)
income_stats = favstats(Price ~ Income, data = income_price)%>%
  arrange(Income)

ggplot(income_price)+
  geom_boxplot(aes(x=Income, y=Price))+
  facet_grid(~Income)
```

(Highway may have effect = uncertainty)

\

#### Claim C:
Gas stations at stoplights charge more
```{r} 
#statistical summary calculation
stoplight_price = gasprices%>%
  select(Price, Stoplight)

ggplot(stoplight_price)+
  geom_boxplot(aes(y=Price))+
  facet_grid(~Stoplight)

favstats(~Price, Stoplight, data = stoplight_price)
mean(stoplight_price$Price ~ stoplight_price$Stoplight)

bootstrap_stoplight = do(10000)*mean(Price ~ Stoplight, data=mosaic::resample(stoplight_price))
confint(bootstrap_stoplight, level = 0.95)
```

\

#### Claim D:
Gas stations with direct highway access charge more
```{r}
highway_price = gasprices%>%
  select(Price, Highway)

ggplot(highway_price)+
  geom_boxplot(aes(y=Price))+
  facet_grid(~Highway)

mean(highway_price$Price ~ highway_price$Highway)

bootstrap_highway = do(10000)*mean(Price ~ Highway, data = mosaic::resample(highway_price))
confint(bootstrap_highway, level = 0.95)
```

\


#### Claim E:
Shell gas stations charge more than any other non-Shell brands 

```{r}
shell_price = gasprices%>%
  select(Price, Name)%>%
  mutate(is_shell = ifelse(Name == 'Shell',
                           yes ='Y', no='N'))
ggplot(shell_price)+
  geom_boxplot(aes(y=Price))+
  facet_grid(~is_shell)

mean(shell_price$Price ~ shell_price$is_shell)

boot_shell = do(10000)*mean(Price ~ is_shell, data = mosaic::resample(shell_price))
confint(boot_shell, level = 0.95)
```


