---
title: "HW8"
author: "Sean Reyes"
date: "2025-04-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
library(tidyverse)
library(boot)

```

#### UTEID: str786
#### Github link:

### Problem 1

Part A:
```{r, echo=FALSE}
# Load the data
creatinine = read.csv("/Users/seanreyes/downloads/creatinine.csv")

# Fit linear regression model
model <- lm(creatclear ~ age, data = creatinine)
summary(model)

# Extract coefficients
intercept <- coef(model)[1]
slope <- coef(model)[2]
```
Using the regression equation:
creatclear = intercept + slope * age, the creatinine clearace rate for 55 year-old is 113.7. This was found because 147.81 represents the clearance rate at birth and the slope of -0.62 is the rate that it deacreases every year which is multiplied by age to result in the predicted clearance rate at age 55.

```{r, echo=FALSE}
pred_55 <- intercept + slope * 55
pred_55
```

\

Part B:
\
Creatinine clearance rate decreases by 0.62 mL minute per year. This was determined form the slope of the model becuase it quantifies the amount of decline that occurs within a year.

\

Part C:
```{r, echo=FALSE}
expected_40 <- intercept + slope * 40
expected_60 <- intercept + slope * 60

residual_40 <- 135 - expected_40
residual_60 <- 112 - expected_60
```
The 40 year olds clearance rate is 11.99 mL/minute above their predicted rate adjusted for their age, but the 60 year olds is only 1.39mL/minute above therefore the 40 year-old's  rate is healthier compared to their age groups normal/expected baseline rate.

\
\

### Problem 2

Part 1:
```{r, echo=FALSE}

covid = read.csv("/Users/seanreyes/downloads/covid.csv")%>%
  filter(deaths > 0)  # Remove days with zero deaths for log transformation

# Define growth model function
fit_growth <- function(data) {
  model <- lm(log(deaths) ~ days_since_first_death, data = data)
  growth_rate <- coef(model)[2]
  doubling_time <- log(2)/growth_rate
  return(c(growth_rate, doubling_time))
}

# Bootstrap function
boot_growth <- function(data, indices) {
  d <- data[indices,]
  return(fit_growth(d))
}

# Analyze Italy
italy <- covid %>% filter(country == "Italy")
italy_boot <- boot(italy, boot_growth, R = 1000)
italy_est <- fit_growth(italy)
italy_growth_ci <- boot.ci(italy_boot, index = 1, type = "perc")
italy_doubling_ci <- boot.ci(italy_boot, index = 2, type = "perc")

# Analyze Spain
spain <- covid %>% filter(country == "Spain")
spain_boot <- boot(spain, boot_growth, R = 1000)
spain_est <- fit_growth(spain)
spain_growth_ci <- boot.ci(spain_boot, index = 1, type = "perc")
spain_doubling_ci <- boot.ci(spain_boot, index = 2, type = "perc")

# Print results for Italy
cat("ITALY RESULTS:\n")
cat(sprintf("Growth rate: %.3f (95%% CI: [%.3f, %.3f])\n", 
            italy_est[1], italy_growth_ci$percent[4], italy_growth_ci$percent[5]))
cat(sprintf("Doubling time: %.1f days (95%% CI: [%.1f, %.1f])\n\n", 
            italy_est[2], italy_doubling_ci$percent[4], italy_doubling_ci$percent[5]))
```

Part 2:
```{r, echo=FALSE}
# Print results for Spain
cat("SPAIN RESULTS:\n")
cat(sprintf("Growth rate: %.3f (95%% CI: [%.3f, %.3f])\n", 
            spain_est[1], spain_growth_ci$percent[4], spain_growth_ci$percent[5]))
cat(sprintf("Doubling time: %.1f days (95%% CI: [%.1f, %.1f])\n\n", 
            spain_est[2], spain_doubling_ci$percent[4], spain_doubling_ci$percent[5]))
```

Part 3:
```{r, echo=FALSE}
# Generate the plot
cat("PLOT GENERATED: See graphical output\n")
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(linewidth = 1) +
  scale_y_continuous(trans = "log10", labels = scales::comma) +
  scale_color_manual(values = c("Italy" = "#1f77b4", "Spain" = "#ff7f0e")) +
  labs(x = "Days since first death", 
       y = "Daily deaths (log scale)",
       title = "COVID-19 Daily Deaths During First Wave",
       color = "Country") +
  theme_minimal(base_size = 12) +
  theme(legend.position = "top")
```



### Problem 3
```{r, echo=FALSE}
# Load data
milk = read.csv("/Users/seanreyes/downloads/milk.csv")

# Fit log-log regression model
model <- lm(log(sales) ~ log(price), data = milk)
beta <- coef(model)[2]

# Bootstrap function for elasticity
boot_elasticity <- function(data, indices) {
  d <- data[indices,]
  model <- lm(log(sales) ~ log(price), data = d)
  return(coef(model)[2])
}

# Run bootstrap (1000 replications)
set.seed(123) # for reproducibility
elasticity_boot <- boot(milk, boot_elasticity, R = 1000)

# Get confidence intervals
beta_ci <- boot.ci(elasticity_boot, type = "perc")

# Print results
cat("PRICE ELASTICITY OF DEMAND FOR MILK:\n")
cat(sprintf("Estimated elasticity (β): %.3f\n", beta))
cat(sprintf("95%% Confidence Interval: [%.3f, %.3f]\n\n", 
            beta_ci$percent[4], beta_ci$percent[5]))

```
To estimate the elasticity a log-log regression model with sales and price was used. After bootstrap smaple the price elasticity was -1.217 with a 95% confidence interval of -1.352 to -1.093. This suggests that the price increase reduces demand by about 1.22%.
